---
title: "Discussion 11 Solution"
output: html_document
---

# Present Group Members

Please **write the names of the present group members** here.

> Present Group Members:

> Absent Group Members:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      error = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tidyverse)
source("../../scripts/ggprob.R")
```

# Preliminaries

-   This file should be in `STAT240/discussion/ds11` on your local computer.

-   This assignment should be completed in your **assigned Canvas groups**. Only one person has to submit on behalf of the group. You may submit whatever you have at the end of discussion section for full credit.

-   This assignment is to be completed in the first 10-20 mins of the discussion session. The rest of the discussion time the students are to work on the group projects with their discussion groups.

## Problems

The questions in this discussion assignment use the chimpanzee data set.
In answering these questions, you may assume the following in models for the data:

1. all trials are independent
2. trials made with a partner have the same chance of being pro-social, even if the identities of the actors or partners differ.

## Problem 1

**Part(a)**

Using the chimpanzee data, combine all of the trials made without a partner.

Count the number of pro-social and selfish choices made.

> Solution

```{r}
chimps = read_csv("../../data/chimpanzee.csv")

prob1 = chimps %>% 
  filter(partner == "none") %>% 
  summarize(prosocial = sum(prosocial),
            selfish = sum(selfish),
            n = prosocial + selfish)

prob1
```

**Part(b)**

Write a model for the number of pro-social choices.

$X$ is the number of pro-social choices made without a partner. $p$ is the probability of making the pro-social choice without a partner present. The number of trials is $n=180$.

$$𝑋\sim \text{Binom}(180,p)$$

From the sample, we get $x=83$.


**Part(c)**

Use the Agresti-Coull method to construct a 95% confidence interval for the probability that a chimpanzee makes the pro-social choice in experimental conditions when there is no partner present.

```{r}
x = 83
n = 180
p_hat_AC = (x+2)/(n+4)
se = sqrt(p_hat_AC*(1-p_hat_AC)/(n+4))
cv = qnorm(0.975)
ci = p_hat_AC + c(-1,1)*cv*se
ci
```


**Part(d)**

Interpret this confidence interval in the context of the experimental setting.

> We are 95% confident that chimpanzees make the pro-social choice with no partner present in the experimental conditions between 38.99 and 53.40 percent of the time.

## Problem 2

Using the same data as in the previous problem,
do the following.

**Part(a)**

State hypotheses for a two-sided hypothesis test that the probability that a chimpanzee makes the pro-social choice over the long run is equal to 0.5.

> $$H_0: p = 0.5 \\ H_{\alpha}: p \neq 0.5$$


**Part(b)**

Pick a test statistic, state the sampling distribution of this statistic if the null hypothesis is true, and compute a p-value for the test.

```{r}
p_value <- 2*pbinom(83, 180, 0.5)
p_value
```

**Part(c)**

Interpret these results in the context of the experiment following the examples from lecture notes.

> The observed data is consistent with the null hypothesis that chimpanzees make the pro-social choice 50% of the time with no partner present ($p=0.33$, two-sided binomial test).

### Please use rest of your time to work on the project proposal in your groups. Come up with an appropriate dataset, a question to ask, and some ideas of how to proceed. 


