---
title: "Section 14: Two-Sample Proportion Inference"
author: "Student's Name Here"
date: "2024-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

```{r}
library(tidyverse)
source("../../scripts/ggprob.R")
```



# CI for p_1 - p_2

Return to the chimpanzee data from last section.

```{r}
chimpanzee <- read_csv("../../data/chimpanzee.csv")

head(chimpanzee)
```

What is the difference in the proportion of prosocial choices by chimpanzee C with versus without a partner?

```{r}
chimpanzee %>% 
  filter(actor == "C") %>% 
  mutate(HasAPartner = (partner != "none")) %>% 
  group_by(HasAPartner) %>% 
  summarize(totalProsocial = sum(prosocial),
            totalSelfish = sum(selfish), 
            n = totalProsocial + totalSelfish, 
            p_hat = totalProsocial/n)
```

Let's build a CI with the Agresti-Coffe adjustment.

```{r}
x1 <- 57
n1 <- 90
x2 <- 17
n2 <- 30

# This code computes the AC Confidence Interval for the true difference in two proportions
n_AC1 <- n1 + 2
n_AC2 <- n2 + 2
p_AC1 <- (x1+1)/n_AC1
p_AC2 <- (x2+1)/n_AC2

pt_est <- p_AC1 - p_AC2

se <- sqrt( p_AC1*(1-p_AC1)/n_AC1 + p_AC2*(1-p_AC2)/n_AC2 )

cv <- qnorm(0.975)

c(pt_est - cv*se, pt_est + cv*se)
```


### Try it out

Build a 95% AC CI for the difference in prosocial behavior for chimpanzee B with and without a partner.

```{r}

```


## Hypothesis testing for p_1 - p_2

Let's look at chimpanzee C again:

```{r}
chimpanzee %>% 
  filter(actor == "C") %>% 
  mutate(HasAPartner = (partner != "none")) %>% 
  group_by(HasAPartner) %>% 
  summarize(totalProsocial = sum(prosocial),
            totalSelfish = sum(selfish), 
            n = totalProsocial + totalSelfish, 
            p_hat = totalProsocial/n)
```


Now, conduct a hypothesis test for the difference in proportions.  The standard error is calculated with an estimated common proportion, p-hat.

```{r}
x1 <- 57
n1 <- 90
x2 <- 17
n2 <- 30

p_hat <- (x1 + x2)/(n1 + n2)

numerator <- (x1/n1) - (x2/n2)  # Observed difference in proportions

denominator <- sqrt(p_hat*(1-p_hat)*(1/n1 + 1/n2))  # Standard error

test_stat <- numerator/denominator
test_stat
```

If H0 is true and p_1 = p_2, then our test statistic came from N(0, 1).  We use the standard normal as our null distribution.  We want to find the area above our test statistic to find evidence that p_1 - p_2 is positive.

```{r}
# Visualization of one-sided p-value
gnorm(col = "black") +
  geom_vline(xintercept = 0.65, color = "red", lty = 2) +
  geom_norm_fill(a = 0.65, fill = "dodgerblue",
                     scale = T)
```

```{r}
pnorm(test_stat, lower.tail = F)
```


### Try it out

Perform a test of *two-sided* hypotheses for p_1 - p_2 != 0, for chimpanzee B.

```{r}

```










