---
title: "week03-sec001"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```



# Sec 004

## Monday 2/3

```{r}
mendota <- read_csv("../../data/lake-mendota-winters-2024.csv")
monona <- read_csv("../../data/lake-monona-winters-2024.csv")
```


```{r}
ggplot(mendota, mapping = aes(x = year1, y = duration)) + geom_point(color = "red") + geom_smooth() + theme_bw()

ggplot() + geom_point(mendota, mapping = aes(x = year1, y = duration), color = "red") + geom_point(monona, mapping = aes(x = year1, y = duration), color = "blue")
```

- First layer:

```{r}
ggplot()

ggplot(mendota, mapping = aes(x = year1, y = duration))
```

- Second layer: using the `geom_functions()`

```{r}
ggplot(mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point() 

ggplot(mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point() + 
  geom_line()
```

# Wednesday 2/5

## Aesthetics:

- variable vs constant:

```{r}
ggplot(mendota, aes(x = year1, y = duration, color = intervals)) +
  geom_point() # variable aes

ggplot(mendota, aes(x = year1, y = duration)) +
  geom_point(mapping = aes(color = intervals)) 

ggplot(mendota, aes(x = year1, y = duration)) +
  geom_point(color = "blue") # constant aesthetic
```

- global vs local aesthetic:

```{r}
# global aesthetic is put inside of ggplot

ggplot(mendota, aes(x = year1, y = duration, color = intervals, size = intervals)) +
  geom_point() +
  geom_line()

# local aesthetic is the ones that go into each geom_functions

ggplot(mendota, aes(x = year1, y = duration)) +
  geom_point(color = "blue", size = 3) +
  geom_line(color = "red", linetype = "dashed")
```

- order of the geom_functions

```{r}
ggplot(mendota, aes(x = year1, y = duration)) +
  geom_point(color = "blue", size = 3) +
  geom_line(color = "red")

ggplot(mendota, aes(x = year1, y = duration)) +
  geom_line(color = "red") +
  geom_point(color = "blue", size = 3)
```

- Alpha:

```{r}
ggplot(mendota, aes(duration)) +
  geom_histogram(
    aes(y = after_stat(density)), # This line shrinks the histogram's height to be on the same scale as geom_density; out of scope of this course, no need to worry about it
    color = "steelblue4",
    fill = "skyblue1"
  ) +
  geom_density(
    fill = "red1",
    alpha = 0.3, # Transparency... the density plot is "on top"
    size = 2
  )  

ggplot(mendota, aes(duration)) +
  geom_density(
    fill = "green",
    size = 2
  ) +
  geom_histogram(
    aes(y = after_stat(density)), # This line shrinks the histogram's height to be on the same scale as geom_density; out of scope of this course, no need to worry about it
    color = "steelblue4",
    fill = "skyblue1",
    alpha = 0.5
  )
```


# Geom_functions:

- `geom_smooth()`: used to create a trendline

```{r}
ggplot(mendota, aes(x = year1, y = duration)) +
  geom_point() +
  geom_smooth()

ggplot(mendota, aes(x = year1, y = duration)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(mendota, aes(x = year1, y = duration)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

- `geom_col()`


```{r}
mendota = mendota %>% 
  mutate(century = as.character(floor(year1/100)+1),
         century = case_when(
           century == "19" ~ "19th",
           century == "20" ~ "20th",
           century == "21" ~ "21st"
         )) 

mendota_summarized = mendota %>% 
  group_by(century) %>% 
  summarize(numYears = n(), avgDuration = mean(duration))

ggplot(mendota_summarized, aes(century, avgDuration)) +
  geom_col()

ggplot(mendota_summarized, aes(century, numYears)) +
  geom_col(color = "blue", # to change the outline of the bars
           fill = "skyblue" # to change the inside of the bars
           )
```

## One Variable Plots

- `geom_histogram()`

```{r}
ggplot(mendota, aes(duration)) +
  geom_histogram(fill = "skyblue", color = "blue")

ggplot(mendota, aes(duration)) +
  geom_histogram(fill = "skyblue", color = "blue", binwidth = 10)

ggplot(mendota, aes(duration)) +
  geom_histogram(fill = "skyblue", 
                 color = "blue", 
                 binwidth = 10,
                 center = 45)

ggplot(mendota, aes(duration)) +
  geom_histogram(fill = "skyblue", 
                 color = "blue", 
                 binwidth = 10,
                 boundary = 40)
```

- `geom_bar()`

```{r}
ggplot(mendota, aes(century)) +
  geom_bar() +
  labs(y = "Number of years in century")
```

- geom_density(): cont var

```{r}
ggplot(mendota, aes(duration)) +
  geom_density()

ggplot(mendota, aes(duration)) +
  geom_density(fill = "red") + geom_histogram(mapping = aes(y = after_stat(density)), color = "black", fill = "white", alpha = 0.7)
```

- geom_boxplot: 

```{r}
ggplot(mendota, aes(duration)) +
  geom_boxplot()
```
with x = duration make a histogram , fill = century

```{r}
ggplot(mendota, aes(duration, fill = century)) +
  geom_histogram(mapping = aes(y = after_stat(density)), color = "black") +
  geom_density()
```

# Faceting

- facet_wrap:

```{r}
ggplot(mendota, aes(duration, fill = as.factor(intervals))) +
  geom_histogram(mapping = aes(y = after_stat(density)), color = "black", bins = 20) +
  geom_density(alpha = 0.7) +
  facet_wrap(facets = vars(century))
```
- facet_grids:

```{r}
ggplot(mendota, aes(duration, fill = century)) +
  geom_density() +
  facet_grid(rows = vars(decade), cols = vars(century))

ggplot(mendota, aes(duration, fill = century)) +
  geom_density() +
  facet_grid(rows = vars(decade))
```


```{r}
ggplot(mendota, aes(duration)) +
  geom_histogram(color = "black", fill = "white") +
  scale_x_continuous(
    breaks = c(30, 60, 90, 120, 150),
    labels = c("1 month", "2 months", "3 months", "4 months", "5 months")
  )
```


