---
title: "week03-sec001"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Sec 001

## Monday 2/3

- `ggplot2` package in `tidyverse` helps us visualize our data

```{r}
mendota <- read_csv("../../data/lake-mendota-winters-2024.csv")
monona <- read_csv("../../data/lake-monona-winters-2024.csv")
```


```{r}
ggplot() + geom_point(mendota, mapping = aes(y = duration, x = year1), color = "red") + geom_point(monona, mapping = aes(y = duration, x = year1), color = "blue")
```

- first layer of the plot

```{r}
ggplot()
ggplot(mendota, mapping = aes(x = duration, y = year1))
```


- second layer: decides the type of graph

```{r}
ggplot(mendota, mapping = aes(x = duration, y = year1)) + geom_point()
```

# Wednesday 2/5

## Variable vs constant aesthetic

```{r}
ggplot(mendota, mapping = aes(x = year1, y = duration)) + geom_point(color = "red", size = 5) # constant aesthetic is that aesrhetic that remains constant for all points or data values in the graph (color, shape, size)

ggplot(mendota, mapping = aes(x = year1, y = duration)) + geom_point(mapping = aes(color = intervals, size = intervals)) # variable aesthetic that is taken from the dataset and varies for the points on the graph
```

## Global vs local aesthetic

```{r}
# global aesthetic is provided inside the ggplot command and affects all the functions following it

ggplot(mendota, mapping = aes(x = year1, y = duration, color = intervals)) +
  geom_line() +
  geom_point()

# local aesthetic is provided to individual geom functions and only affects the respective geom function
ggplot(mendota, mapping = aes(x = year1, y = duration)) +
  geom_line(color = "red") +
  geom_point(color = "blue", size = 5)


ggplot(mendota, mapping = aes(x = year1, y = duration)) +
  geom_point(color = "blue", size = 5) +
   geom_line(color = "red") 
```


```{r}
# global vs local aesthetic

ggplot(mendota, mapping = aes(x = year1, y = duration)) +
  geom_line()

ggplot() +
  geom_line(data = mendota, mapping = aes(x = year1, y = duration), color = "red") +
  geom_line(data = monona, mapping = aes(x = year1, y = duration), color = "blue")
```

- Understanding `alpha`

```{r}
ggplot(mendota, aes(duration)) +
  geom_histogram(
    aes(y = after_stat(density)), # This line shrinks the histogram's height to be on the same scale as geom_density; out of scope of this course, no need to worry about it
    color = "steelblue4",
    fill = "skyblue1"
  ) +
  geom_density(
    fill = "red1",
    alpha = 0.3, # Transparency... the density plot is "on top"
    size = 2
  )  

ggplot(mendota, aes(duration)) +
  geom_histogram(
    aes(y = after_stat(density)), # This line shrinks the histogram's height to be on the same scale as geom_density; out of scope of this course, no need to worry about it
    color = "steelblue4",
    fill = "skyblue1"
  ) +
  geom_density(
    fill = "green",
    alpha = 1, # Transparency... the density plot is "on top"
    size = 1
  ) 

```



```{r}
ggplot(mendota, mapping = aes(x = year1, y = duration)) + geom_point() + geom_smooth()

ggplot(mendota, mapping = aes(x = year1, y = duration)) + geom_point() + geom_smooth(method = "lm")


ggplot(mendota, mapping = aes(x = year1, y = duration)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```


```{r}
mendota = mendota %>% 
  mutate(century = as.character(floor(year1/100)+1),
         century = case_when(
           century == "19" ~ "19th",
           century == "20" ~ "20th",
           century == "21" ~ "21st"
         ))

mendota_summarized = mendota %>% 
  group_by(century) %>% 
  summarize(numYears = n(), avgDuration = mean(duration)) 
  

ggplot(mendota_summarized, aes(century, numYears)) +
  geom_col(
    color = "steelblue4",
    fill = "skyblue1"
  )

ggplot(mendota_summarized, aes(century, avgDuration)) +
  geom_col(
    color = "steelblue4",
    fill = "skyblue1"
  )
```

# Friday 2/7

## One variables:

- geom_histogram: continuous variables

```{r}
ggplot(mendota, mapping = aes(x = duration)) + 
  geom_histogram(color = "blue",
                 fill = "skyblue",
                 bins = 20,
                 boundary = 40)
```


- geom_bar: categorical variables

```{r}
ggplot(mendota, aes(century)) +
  geom_bar(fill = "skyblue")
```

- geom_density: 

```{r}
ggplot(mendota, aes(duration)) +
  geom_density()
```

- geom_boxplot:

```{r}
ggplot(mendota, aes(duration)) +
  geom_boxplot()
```

```{r}
ggplot(mendota, aes(duration, fill = century)) +
  geom_histogram(aes(y = after_stat(density))) +
  geom_density(alpha = 0.9)
```

## Faceting

- facet wrap

```{r}
ggplot(mendota, aes(x = duration, fill = century)) +
  geom_density() +
  facet_wrap(facets = vars(century))
```

- facet_grid

```{r}
ggplot(mendota, aes(x = duration, fill = century)) +
  geom_density() +
  facet_grid(rows = vars(decade), cols = vars(century))

ggplot(mendota, aes(x = duration, fill = century)) +
  geom_density() +
  facet_grid(rows = vars(decade))
```

```{r}
ggplot(mendota, aes(duration))+
  geom_histogram() +
  scale_x_continuous(
    breaks = c(30, 60, 90, 120),
    labels = c("1 month", "2 months", "3 months", "4 months")
  )
```

# Monday 2/10

- PLot labels: `labs()`

```{r}
ggplot(mendota, aes(duration))+
  geom_histogram(color = "black", fill = "white") +
  scale_x_continuous(  # will change the scale 
    breaks = c(30, 60, 90, 120),
    labels = c("1 month", "2 months", "3 months", "4 months")
  ) + 
  labs(  # will give labels
    title = "Histogram for the duration of lake freezing",
    x = "duration (in days)",
    y = "frequency"
  ) +
  theme_minimal() # will change theme

x = mean(mendota$duration)

ggplot(mendota, aes(duration))+
  geom_histogram(color = "black", fill = "white") +
  scale_x_continuous(  # will change the scale 
    breaks = c(30, 60, 90, 120),
    labels = c("1 month", "2 months", "3 months", "4 months")
  ) + 
  labs(  # will give labels
    title = "Histogram for the duration of lake freezing",
    x = "duration (in days)",
    y = "frequency"
  ) +
  theme_minimal() +
  geom_vline(xintercept = x, color = "red", linetype = "dashed")
```


```{r}
ggplot(mendota, aes(year1, duration, color = intervals)) +
  geom_point()

ggplot(mendota, aes(year1, duration, color = as.factor(intervals))) +
  geom_point() +
  labs(color = "intervals")
```













