```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = TRUE,
                      error = TRUE, fig.height = 4)
library(tidyverse)
source("../../scripts/ggprob.R")

```
## Bob Ross The Joy of Painting Case Study (Review)

### Read elements data

```{r}
all_elements = read_csv("../../data/elements-by-episode.csv")

all_elements
```

> Are the titles unique?

```{r}
all_elements %>% 
  count(TITLE) %>% 
  filter(n > 1)
```

- No, two titles were repeated twice each.

> Find episodes with a repeated title

```{r}
repeated_titles = all_elements %>% 
  count(TITLE) %>% 
  filter(n > 1) %>% 
  pull(TITLE)

all_elements %>%
  filter(TITLE %in% repeated_titles) %>% 
  select(EPISODE, TITLE)
```

#### semi-join() solution

```{r}
repeated_titles_df = all_elements %>% 
  count(TITLE) %>% 
  filter(n > 1)

all_elements %>% 
  semi_join(repeated_titles_df, join_by(TITLE)) %>% 
  select(EPISODE, TITLE)
```

> Create a data frame with only Bob Ross hosted paintings and no host or frame variables.

```{r}
painting_elements = all_elements %>% 
  filter(GUEST == 0) %>% 
  select(-GUEST, -DIANE_ANDRE, -STEVE_ROSS, -FRAMED,
         -contains("_FRAME"))

painting_elements
```

> Write code which counts the number of times each feature appears.

### Approach #1 across()

```{r}
feature_counts = painting_elements %>% 
 # summarize(AURORA_BOREALIS = sum(AURORA_BOREALIS))
  summarize(across(AURORA_BOREALIS:WINTER, sum ))

feature_counts = painting_elements %>% 
 # summarize(AURORA_BOREALIS = sum(AURORA_BOREALIS))
  summarize(across(AURORA_BOREALIS:WINTER, sum))

### Example with more than one function
#  summarize(across(AURORA_BOREALIS:WINTER, list(sum = sum, mean = mean)))

feature_counts
```


### Approach #2 pivot_longer()

- Get counts and percentages
- Arrange in order from high to low

```{r}
painting_elements_long = painting_elements %>% 
  pivot_longer(AURORA_BOREALIS:WINTER,
               names_to = "feature",
               values_to = "value")

feature_counts = painting_elements_long %>% 
  group_by(feature) %>% 
  summarize(n = sum(value),
            pct = round(100*mean(value))) %>% 
  arrange(desc(n))

feature_counts
```


### Visualize counts

- basic bar plot

```{r}
ggplot(feature_counts, aes(x = pct, y = feature)) +
  geom_col()
```

- change the order of the features

```{r}
ggplot(feature_counts, aes(x = pct,
                           y = reorder(feature, pct))) +
  geom_col()
```

- add better labels
- custom theme (complicated)

```{r}
ggplot(feature_counts, aes(x = pct,
                           y = reorder(feature, pct))) +
  geom_col() +
  labs(
    title = "The Paintings of Bob Ross",
    subtitle = "Percentage containing each element",
    x = "",
    y = ""
  ) +
  ### lots of control, hard to learn
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    plot.background = element_rect(fill = "gray96", color = "gray96"),
    plot.title.position = "plot",
    plot.title = element_text(face = "bold")
  )
```

- filter for pct >= 2
- add % numbers

```{r}
feature_counts %>% 
  filter(pct > 1) %>% 
ggplot( aes(x = pct,
                           y = reorder(feature, pct))) +
  geom_col() +
  ### See solution in notes for a better way!
  geom_text(aes(x = pct + 5, label = pct)) +
  labs(
    title = "The Paintings of Bob Ross",
    subtitle = "Percentage containing each element",
    x = "",
    y = ""
  ) +
  ### lots of control, hard to learn
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    plot.background = element_rect(fill = "gray96", color = "gray96"),
    plot.title.position = "plot",
    plot.title = element_text(face = "bold")
  )
```

- Even more can be done;;; see the notes

### Painting Colors

- New data with the colors used in each episode

#### Read the data

```{r}
painting_colors = read_csv("../../data/bob_ross_paintings.csv")
```

### Visualize num_colors

- histogram
  - always specify `binwidth` and either `boundary` or `center`
  - also good to specify `color` (edge color) and `fill` (interior color)

```{r}
ggplot(painting_colors, aes(x = num_colors)) +
  geom_histogram(center = 1, binwidth = 1,
                 color = "black", fill = "hotpink")
  
```

### Join data sets

- Add `season` and `episode` columns to `painting_elements`
- Then, `left_join()` with `painting_colors`

```{r}
## Add season and episode using str_sub()

painting_elements = painting_elements %>% 
  mutate(season = as.numeric(str_sub(EPISODE, 2, 3)),
         episode = as.numeric(str_sub(EPISODE, 5, 6))) %>% 
  relocate(season, episode, .before = EPISODE)

temp_colors = painting_colors %>% 
  select(season, episode, painting_title, num_colors,
         Black_Gesso:Alizarin_Crimson)

ross = painting_elements %>% 
  left_join(temp_colors, join_by(season, episode)) %>% 
  relocate(painting_title, num_colors,
           .after = TITLE) %>% 
  select(-TITLE) %>% 
  rename(ID = EPISODE)
```

### Inference

- 95% CI for paintings with a man-made structure

```{r}
prob1 = ross %>% 
  summarize(n = n(),
            ntilde = n+4,
            x = sum(STRUCTURE),
            xtilde = x+2,
            phat = x/n,
            ptilde = xtilde/ntilde,
            se = sqrt(ptilde * (1 - ptilde) / ntilde),
            z_crit = qnorm(0.975),
            moe = z_crit * se,
            low = ptilde - moe,
            high = ptilde + moe)

prob1 %>% 
  print(width = Inf)

prob1 %>% 
  select(low, high) %>% 
  as.data.frame()
  
```

- 95% HT for $p$, the hypothetical population proportion of paintings with a man-made structure.

$H_0: p =  0.2$  
$H_a: p \neq 0.2$

If $H_0$ is true, then $X \sim \text{Binomial}(381, 0.2)$.

The observed statistic is $x = 84$.

Compare to the binomial distribution:

$\mathsf{E}(X) = 381 \times 0.2 \doteq `r 381*0.2`$

```{r}
expected = 381*0.2
expected
```

$\mathsf{SE}(X) = \sqrt{ 381 \times 0.2 \times 0.8} \doteq `r round(sqrt(381*0.2*0.8),1)`$

```{r}
se = sqrt(381*0.2*0.8)
se
```

- The observed $x = 84$ is about 1 SE above the mean, which is not unusual.

- Calculate the two-sided p-value by finding $P(X \ge 84)$ for the right tail and $P(X \le 68)$ for the left tail.
    - The notes has a likelihood-based method to determine the cutoff of the left tail.
    - A short cut is $x=84$ is 7.8 above the mean, the mean 76.2 minus 7.8 is 68.4, so any result 68 or lower is more extreme than the observed $x=84$.

- Calculate p-value from direct binomial probability calculation

```{r}
p_value = 1 - pbinom(83, 381, 0.2) + pbinom(68, 381, 0.2)
p_value
```

> The data is consistent with the null hypothesis $p=0.2$, that the long-run proporiton of times Bob Ross includes a man-made structure in his TJOP paintings equals 0.2, (p-value = 0.33, two-sided binomial test for $p$).

OR

> There is a lack of sufficient statistical evidence to conclude that
the long-run proporiton of times Bob Ross includes a man-made structure in his TJOP paintings is not equal to 0.2, $p \neq 0.2$, (p-value = 0.33, two-sided binomial test for $p$).


### Differ from posted notes

- Separate populations by Phthalo Green or not.
- Confidence interval for difference in proportions with a man-made structure.

$p_1$ ~ P(man-made structure | Phthalo Green)
$p_2$ ~ P(man-made structure | not Phthalo Green)

95% CI for $p_1 - p_2$

```{r}
prob2_sum = ross %>% 
  group_by(Phthalo_Green) %>% 
  summarize(x = sum(STRUCTURE),
            n = n(),
            phat = x/n,
            xtilde = x+1,
            ntilde = n+2,
            ptilde = xtilde / ntilde,
            se = sqrt(ptilde*(1-ptilde)/ntilde))

prob2_sum %>% 
  print(width = Inf)

prob2_ci = prob2_sum %>% 
  summarize(est = diff(ptilde),
            z_crit = qnorm(0.975),
            se = sqrt(sum(se^2)),
            moe = z_crit * se,
            low = est - moe,
            high = est + moe)

prob2_ci
```

> We are 95% confident that the difference in population proportions of Bob Ross paintings with a man-made structure between paintings with and without phthalo green is between -16.8% and 0.08%.

#### Hypothesis Test

$H_0: p_1 = p_2$    
$H_a: p_1 \neq p_2$

```{r}
prob2_ht = prob2_sum %>% 
  summarize(est = diff(phat),
            pbar = sum(x) / sum(n),
            se = sqrt(sum(pbar*(1-pbar)/n)),
            z_test = est / se,
            p_value = 2*pnorm(-abs(z_test)))

prob2_ht
```


> There is insufficient evidence to conclude there is a difference in the population proprtions of Bob Ross paintings with a man-made structure between paintings with and without phthalo green (p-value = 0.079, two-sided z-test for proportions)


