```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = TRUE,
                      error = TRUE, fig.height = 4)
library(tidyverse)
source("../../scripts/ggprob.R")

```
## Bob Ross The Joy of Painting Case Study (Review)

### Read elements data

```{r}
all_elements = read_csv("../../data/elements-by-episode.csv")

all_elements
```

> Are the titles unique?

```{r}
all_elements %>% 
  count(TITLE) %>% 
  filter(n > 1)
```

- No, two titles were repeated twice each.

> Find episodes with a repeated title

```{r}
repeated_titles = all_elements %>% 
  count(TITLE) %>% 
  filter(n > 1) %>% 
  pull(TITLE)

all_elements %>%
  filter(TITLE %in% repeated_titles) %>% 
  select(EPISODE, TITLE)
```

#### semi-join() solution

```{r}
repeated_titles_df = all_elements %>% 
  count(TITLE) %>% 
  filter(n > 1)

all_elements %>% 
  semi_join(repeated_titles_df, join_by(TITLE)) %>% 
  select(EPISODE, TITLE)
```

> Create a data frame with only Bob Ross hosted paintings and no host or frame variables.

```{r}
painting_elements = all_elements %>% 
  filter(GUEST == 0) %>% 
  select(-GUEST, -DIANE_ANDRE, -STEVE_ROSS, -FRAMED,
         -contains("_FRAME"))

painting_elements
```

> Write code which counts the number of times each feature appears.

### Approach #1 across()

```{r}
feature_counts = painting_elements %>% 
 # summarize(AURORA_BOREALIS = sum(AURORA_BOREALIS))
  summarize(across(AURORA_BOREALIS:WINTER, sum ))

feature_counts = painting_elements %>% 
 # summarize(AURORA_BOREALIS = sum(AURORA_BOREALIS))
  summarize(across(AURORA_BOREALIS:WINTER, sum))

### Example with more than one function
#  summarize(across(AURORA_BOREALIS:WINTER, list(sum = sum, mean = mean)))

feature_counts
```


### Approach #2 pivot_longer()

- Get counts and percentages
- Arrange in order from high to low

```{r}
painting_elements_long = painting_elements %>% 
  pivot_longer(AURORA_BOREALIS:WINTER,
               names_to = "feature",
               values_to = "value")

feature_counts = painting_elements_long %>% 
  group_by(feature) %>% 
  summarize(n = sum(value),
            pct = round(100*mean(value))) %>% 
  arrange(desc(n))

feature_counts
```


### Visualize counts

- basic bar plot

```{r}
ggplot(feature_counts, aes(x = pct, y = feature)) +
  geom_col()
```

- change the order of the features

```{r}
ggplot(feature_counts, aes(x = pct,
                           y = reorder(feature, pct))) +
  geom_col()
```

- add better labels
- custom theme (complicated)

```{r}
ggplot(feature_counts, aes(x = pct,
                           y = reorder(feature, pct))) +
  geom_col() +
  labs(
    title = "The Paintings of Bob Ross",
    subtitle = "Percentage containing each element",
    x = "",
    y = ""
  ) +
  ### lots of control, hard to learn
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    plot.background = element_rect(fill = "gray96", color = "gray96"),
    plot.title.position = "plot",
    plot.title = element_text(face = "bold")
  )
```

- filter for pct >= 2
- add % numbers

```{r}
feature_counts %>% 
  filter(pct > 1) %>% 
ggplot( aes(x = pct,
                           y = reorder(feature, pct))) +
  geom_col() +
  ### See solution in notes for a better way!
  geom_text(aes(x = pct + 5, label = pct)) +
  labs(
    title = "The Paintings of Bob Ross",
    subtitle = "Percentage containing each element",
    x = "",
    y = ""
  ) +
  ### lots of control, hard to learn
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    plot.background = element_rect(fill = "gray96", color = "gray96"),
    plot.title.position = "plot",
    plot.title = element_text(face = "bold")
  )
```

- Even more can be done;;; see the notes

### Painting Colors

- New data with the colors used in each episode

#### Read the data

```{r}
painting_colors = read_csv("../../data/bob_ross_paintings.csv")
```

### Visualize num_colors

- histogram
  - always specify `binwidt`h and either `boundary` or `center`
  - also good to specify `color` (edge color) and `fill` (interior color)

```{r}
ggplot(painting_colors, aes(x = num_colors)) +
  geom_histogram(center = 1, binwidth = 1,
                 color = "black", fill = "hotpink")
  
```

... to be continued

