---
title: "Section 13: Proportion Inference"
author: "Student's Name Here"
date: "2024-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

```{r}
library(tidyverse)
source("../../scripts/ggprob.R")
```

## Proportion sampling distribution


This lecture will use the data in `chimpanzee.csv`. The data is based on an experiment conducted by scientists at Emory University in 2011; though those scientists did not make their true data public, we have created a "mock" dataset which gives the same results and summary values as their data.

* Two chimpanzees were put in adjacent enclosures with the ability to see and communicate with each other.

* One chimpanzee, the **actor**, was given the choice to pick from a bucket of tokens which were two different colors.

* The actor had previously been taught that one color would lead to the human researcher to give them (just the actor) a treat. We call this the **selfish** choice.

* The other color would lead to the human researcher giving **both the actor and the partner** a treat. We call this the "**prosocial**" choice.

* The **actor** was given this choice 30 times, and the number of **prosocial** and **selfish** choices were recorded.

* The researchers performed this experiment with many combinations of actors, partners, and colors; including observing their behavior when the actor had **no partner**.

---

* Our "mock" dataset for this can be found in `chimpanzee.csv`.  Each row is a combination of actor and partner

```{r}
chimpanzee <- read_csv("../../data/chimpanzee.csv")

head(chimpanzee)
```

Let's exploree the data.

```{r}
sum1 <- chimpanzee %>% 
  mutate(session_type = case_when(
    partner == "none" ~ "no partner",
    TRUE ~ "partner"
  )) %>% 
  group_by(actor, session_type) %>% 
  
  summarize(prosocial = sum(prosocial),
            selfish = sum(selfish),
            n = prosocial + selfish,
            pct_prosocial = 100*prosocial/n)
sum1

sum1_wide <- sum1 %>% 
  select(actor, session_type, pct_prosocial) %>% 
  pivot_wider(names_from = session_type,
              values_from = pct_prosocial)

sum1_wide
```

```{r}
ggplot(sum1, aes(x = actor, y = pct_prosocial,
                 fill = session_type)) +
  geom_col(color = "black",
           position = position_dodge2(preserve = "single")) +
  #scale_y_continuous(labels = percent_format(scale = 1)) +
  labs(
    x = "Chimpanzee",
    y = "Pro-social Choice Probability",
    title = "Chimpanzee Pro-social Choice Comparison",
    subtitle = "With and Without Partners",
    fill = "Session"
  ) +
  theme_minimal()
```



```{r}
## simulation parameters
B <- 100000
n <- 90
x <- 60
p <- x/n

## do the simulation
sim <- tibble(
  x_star = rbinom(B, n, p),
  p_hat = x_star / n)
```

```{r}
## Graphical summary
ggplot(sim, aes(x = p_hat, y=after_stat(density))) +
  geom_histogram(binwidth = 1/90,
                 color = "black", fill = "dodgerblue") +
  xlab("sample proportion (p-hat)") +
  theme_minimal() +
  # Theoretical sampling distribution
  geom_norm_density(mu=p, sigma=sqrt(p*(1-p)/n),color="blue", size=2) 

## Numeric summary
sim_summary <- sim %>% 
  summarize(mean = mean(p_hat),  ## this should be very close to 0.667
            sd = sd(p_hat)) ## this should be very close to 0.0497
sim_summary
```

The sampling distribution of p-hat has a normal shape, and its mean is very close to $p = 0.667$.


### Try it out

Chimpanzee D made 8 prosocial choices and 22 selfish choices with no partner. Recreate the simulation for $\hat{p}_{D, nopartner}$.  What do you notice about the shape of the histogram?

```{r}

```


## Confidence interval for p

For 90% confidence (alpha = 0.1), use the 95th or 5th percentile.

```{r}
qnorm(0.05)
qnorm(0.95)
```

A 90% Wald CI for $p_{A, partner}$ is

```{r}
n <- 90
x <- 60
p_hat <- x/n

se <- sqrt(p_hat * (1 - p_hat) / n)

cv <- qnorm(0.95)


c(p_hat - cv*se, p_hat + cv*se)
```

Now, let's make the Agresti-Coull adjustment.  We add two successes and failures to our data.

```{r}
n <- 90
x <- 60

n_ac <- n + 4
p_hat_ac <- (x + 2) / (n_ac)

se <- sqrt(p_hat_ac * (1 - p_hat_ac) / n_ac)

cv <- qnorm(0.95)


c(p_hat_ac - cv*se, p_hat_ac + cv*se)
```


## Hypothesis test for p

We want to test whether chimpanzee A makes prosocial choices with a rate greater than 0.5. The null distribution is Binom(90, 0.5) and the test statistic is 60.

```{r}
gbinom(90, 0.5, col = "black") +
  geom_binom_density(90, 0.5, a = 60, lwd = 3, col = "dodgerblue")

gbinom(90, 0.5, col = "black") +
  geom_binom_density(90, 0.5, a = 60, lwd = 3, col = "dodgerblue") +
  ylim(0, 0.01) + xlim(55, 90) +
  annotate("text", x = 65, y = 0.0025, label = "zoom... and enhance!")
```

The p-value is $P(Binom(90, 0.5) \ge 60)$

```{r}
1 - pbinom(59, 90, 0.5)
```


### Try it out

```{r}
sum1_partner <- sum1 %>%
  filter(session_type == "partner")

# Overall prosocial rate with a partner
sum(sum1_partner$prosocial) / sum(sum1_partner$n)
```

Chimpanzee F has the lowest observed prosocial-partner rate of 47/90.  Is the true rate significantly *less* than the average of 0.59?


```{r}

```


### Psychic example

A psychic claims to be able to guess a card suit at a better rate than p = 0.25.  They get 57 guesses correct out of 200.  The probability of 57, under the null, is 0.033.  Which outcomes are less likely?

```{r}
gbinom(200, 0.25, scale = TRUE, col = "black") +
  geom_vline(xintercept = 57, color = "red", lty = 2) +
  geom_hline(yintercept = dbinom(57, 200, 0.25), color = "red", lty = 2)
```

The outcomes 42 and below are less likely (more extreme, more evidence against H0) as 57.

```{r}
dbinom(57, 200, 0.25)

tibble(
  x = 38:45,
  prob = dbinom(x, 200, 0.25),
  lessThanP57 = prob < dbinom(57, 200, 0.25)
)
```


For a one-sided p-value, find the probability of X >= 57.

```{r}
1 - pbinom(57 - 1, 200, 0.25)
```

```{r}
# Visualization of one-sided p-value
gbinom(200, 0.25, scale = TRUE, col = "black") +
  geom_vline(xintercept = 57, color = "red", lty = 2) +
  geom_binom_density(200, 0.25, a = 57, col = "dodgerblue",
                     scale = T)
```


What if our hypotheses were two-sided?  The p-value includes the outcomes greater than or equal to 57 AND those less than or equal to 42.

```{r}
pbinom(42, 200, 0.25) + (1 - pbinom(57 - 1, 200, 0.25))
```

```{r}
# Visualization of two-sided p-value
gbinom(200, 0.25, scale = TRUE, col = "black") +
  geom_vline(xintercept = 57, color = "red", lty = 2) +
  geom_binom_density(200, 0.25, a = 57, col = "dodgerblue",
                     scale = T) +
    geom_binom_density(200, 0.25, b = 42, col = "dodgerblue",
                     scale = T)
```


# Normal approximation

Let's repeat the two-sided test in the psychic example with a Z test.  Our null distirbution is N(0, 1) and our test statistic is 1.143.

```{r}
# Visualization of two-sided p-value
gnorm(col = "black") +
  geom_vline(xintercept = 1.143, color = "red", lty = 2) +
  geom_norm_fill(a = 1.143, fill = "dodgerblue",
                     scale = T) +
    geom_norm_fill(b = -1.143, fill = "dodgerblue",
                     scale = T)
```

P-value calculation:

```{r}
pnorm(-1.143) + pnorm(1.143, lower.tail = F)
# or
2*pnorm(1.143, lower.tail = F)
```

