---
title: "week04_in_class_noted"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(tidyverse) # library(dplyr)
library(magrittr)
```

## Pipe Operator

- `%>%` is known as the pipe operator

- MAC: `cmd+shift+M`
- Windows: `ctrl+shift+M`

- x %>% f() -> f(x)
- x %>% f() %>% g() -> f(x) %>% g() -> g(f(x))
- x %>% f(y1, y2) %>% g(z) -> f(x, y1, y2) %>% g(z) -> g(f(x, y1, y2), z)

Example:

$$\sqrt{\frac{(3+4)^2}{7} } $$

```{r}
sqrt(divide_by(raise_to_power(add(3,4),2), 7))

3 %>% add(4) %>% raise_to_power(2) %>% divide_by(7) %>% sqrt()
```

## Loading the dataset

```{r}
grocery_list <- read_csv("../../data/grocery-list.csv")
grocery_price <- read_csv("../../data/grocery-prices.csv")
```


```{r}
produce <- grocery_list %>% left_join(grocery_price, by = "item")

produce
```

# Wednesday 2/12

## Column editors

- `mutate()`: adds a column 

```{r}
produce1 = produce %>% mutate(cost = quantity*price, tax = 0.05*cost, total_cost = cost+tax, index = 1:10)

produce1 %>% select(item, total_cost)

produce1 %>% select(- c(quantity:tax), -index)
```

- `select()`: selecting specific columns from the dataset

```{r}
produce %>% 
  mutate(cost = quantity*price, tax = 0.05*cost, total_cost = cost+tax, index = 1:10) %>% 
  select(item, price, total_cost)

produce1 %>% 
  select(item, price, total_cost)


```

- `relocate()`: relocates columns in your dataset

```{r}
produce1

produce1 %>% relocate(index, quantity)

produce1 %>% relocate(index) %>% relocate(cost, tax, total_cost, .before = type)
```
- `rename()`: renaming of columns. `new name = old name`

```{r}
produce1 %>% rename(sales_tax = tax, n = quantity)

produce1 %>% rename(quantity = n) # gives error if you put old name = new name
```

```{r}
produce %>% mutate(index = 1:10, cost = quantity*price, tax = 0.05*cost, total_cost = cost+tax) %>% 
  relocate(index) %>% 
  select(-type)
```

## Rows editors

- `arrange()`: arranges columns by default in the ascending order. use the function `desc()` or the `-` sign (just for the numeric variable) to arrange in descending order


```{r}
produce %>% arrange(price)
produce %>% arrange(-price) #arrange(desc(price))
produce %>% arrange(desc(item))
produce %>% arrange(-item) # gives errors since it is a character column
```

- `filter()`:filters rows depending on the condition provided, logical operators are used to provide the condition or logical argument

```{r}
produce1 %>% filter(cost>7)
produce1 %>% filter(quantity > 2 & total_cost < 4) 
```

```{r}
produce %>% filter(type == "fruit")

produce %>% filter(price+5)
```

- `drop_na()`

```{r}
starwars

starwars %>% drop_na(hair_color)

starwars %>% drop_na()

starwars %>% filter(!is.na(mass), !is.na(hair_color))
```


## summarize()

```{r}
sqrt(1:10)

sum(1:10)
mean(1:10)
median(1:10)
```

```{r}
produce1 %>% summarize(total_spent = sum(total_cost), avg_spent = mean(total_cost))

produce1 %>% filter(type == "fruit") %>% summarize(total_spent = sum(total_cost))
```
## group_by()

```{r}
produce1

produce1 %>% group_by(type) %>% summarize(total_spent = sum(total_cost))

produce1 %>% group_by(price < 1) %>% summarize(total_spent = sum(total_cost))
```

## case_when

```{r}
produce2 <- produce1 %>% mutate(categories = case_when(
  price <= 1 ~ "price less than $1",
  price >1 & price <= 2 ~ "price between $1 and $2",
  price >2 ~ "price more than $2"
))

produce2 %>% group_by(categories) %>% summarize(total_spent = sum(total_cost))
```

```{r}
produce2 %>% summarize(num_items = n())
produce2 %>% group_by(categories) %>% summarize(num_items = n())
produce2 %>% group_by(categories) %>% count()

produce2 %>% group_by(categories) %>% summarize(num_items = n(), total_quantity = sum(quantity), total_spent = sum(total_cost))
```

## slice_min and slice_max

```{r}
produce1 %>% slice_max(total_cost)

produce1 %>% slice_max(total_cost, n=3)

produce1 %>% group_by(type) %>% slice_max(total_cost, n=2)
```



```{r}
produce1 %>% group_by(type) %>% summarize(max = max(total_cost), min = min(total_cost))
```
## group_by with mutate

```{r}
produce2 %>% group_by(categories) %>% summarize(total_spent = sum(total_cost))

produce2 %>% group_by(categories) %>% mutate(total_spent = sum(total_cost))

produce2 %>% mutate(total_spent = sum(total_cost))
```

```{r}
t.test(x, y, paired = T)
```




