---
title: "week04_in_class_notes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, warning = TRUE)
library(tidyverse)
library(magrittr)
```

## Pipe operator:

-   `%>%` is called the pipe operator
    -   Mac: `cmd+shift+M`
    -   Window: `ctrl+shift+M`
-   Working
    -   x %\>% f() -\> f(x)
    -   x %\>% f() %\>% g() -\> f(x) %\>% g() -\> g(f(x))
    -   x %\>% f(y) %\>% g(z) -\> f(x,y) %\>% g(z) -\> g(f(x,y),z)

Example: Calculating this using the pipe operator

$$\sqrt{\frac{(3+4)^2}{3}}$$

```{r}
3 %>% add(4) %>% raise_to_power(2) %>% divide_by(3) %>% sqrt()

sqrt(divide_by(raise_to_power(add(3, 4), 2), 3))
```

## Read in the datasets:

```{r}
grocery_list <- read_csv("../../data/grocery-list.csv")
grocery_price <- read_csv("../../data/grocery-prices.csv")
```
```{r}
produce <- grocery_list %>% left_join(grocery_price, by = "item")

produce
```

# Wednesday 2/12

## Column editors

- `mutate()`: adds a column 

```{r}
produce1 = produce %>% mutate(cost = quantity*price, tax = 0.05*cost, total_cost = cost+tax, index = 1:10)
```

- `select()`: selecting specific columns from the dataset

```{r}
produce %>% 
  mutate(cost = quantity*price, tax = 0.05*cost, total_cost = cost+tax, index = 1:10) %>% 
  select(item, price, total_cost)

produce1 %>% 
  select(item, price, total_cost)
```

- `relocate()`: relocates columns in your dataset

```{r}
produce1

produce1 %>% relocate(index, quantity)

produce1 %>% relocate(index) %>% relocate(cost, tax, total_cost, .before = type)
```
- `rename()`: renaming of columns. `new name = old name`

```{r}
produce1 %>% rename(sales_tax = tax, n = quantity)

produce1 %>% rename(quantity = n) # gives error if you put old name = new name
```

```{r}
produce %>% mutate(index = 1:10, cost = quantity*price, tax = 0.05*cost, total_cost = cost+tax) %>% 
  relocate(index) %>% 
  select(-type)
```

## Rows editors

- `arrange()`: arranges columns by default in the ascending order. use the function `desc()` or the `-` sign (just for the numeric variable) to arrange in descending order


```{r}
produce %>% arrange(price)
produce %>% arrange(-price) #arrange(desc(price))
produce %>% arrange(desc(item))
produce %>% arrange(-item) # gives errors since it is a character column
```

- `filter()`:filters rows depending on the condition provided, logical operators are used to provide the condition or logical argument

```{r}
produce1 %>% filter(cost>7)
produce1 %>% filter(quantity > 2 & total_cost < 4) 
```

```{r}
produce %>% filter(type == "fruit")

produce %>% filter(price+5)
```

- `drop_na()`

```{r}
starwars

starwars %>% drop_na(hair_color)

starwars %>% drop_na()

starwars %>% filter(!is.na(mass), !is.na(hair_color))
```


## Exercise

“Show me just the fruits.”

```{r}
produce %>% filter(type == "fruit")
```


“Can you put whether it’s a fruit or vegetable at the beginning?”

```{r}
produce %>% relocate(type)
```



“I only care about the food name and the total cost, can you get rid of everything else?”


```{r}
produce %>% mutate(total_cost = (1+0.05)*quantity*price) %>% select(item, total_cost)

produce %>% mutate(total_cost = (1+0.05)*quantity*price) %>% select(-c(quantity:type))
```


“Can you put the foods in the opposite alphabetical order so apple is last?”

```{r}
produce %>% arrange(desc(item))
```


“Can you give each food a sale price that’s half off the normal price?”

```{r}
produce %>% mutate(sale_price = 0.5*price)
```

```{r}
sum(1:10)
mean(1:10)

sqrt(1:10)
```


```{r}
produce1 %>% summarize(Total_spent = sum(total_cost), avg = mean(total_cost))

produce1 %>% filter(type == "fruit") %>% summarize(total_cost_fruit = sum(total_cost))
```


```{r}
produce1
produce1 %>% group_by(type) %>% summarise(avg_spent = mean(total_cost))

produce1 %>% group_by(type) %>% mutate(avg_spent = mean(total_cost)) %>% relocate(avg_spent)

produce1 %>% group_by(type) %>% slice_max(total_cost, n = 3)
produce1 %>% group_by(type) %>% slice_min(total_cost, n = 3)
```


```{r}
produce1 %>% summarise(avg_spent = mean(total_cost), num_elements = n())

produce1 %>% group_by(type) %>% summarise(avg_spent = mean(total_cost), num_elements = n())

produce1 %>% group_by(type) %>% summarise(num_elements = n())

produce1 %>% group_by(type) %>% count()
```

```{r}
produce1 %>% group_by(price < 1) %>% summarise(total_spent = sum(total_cost))
```

```{r}
produce2 <- produce1 %>% mutate(group = case_when(
  price <= 1 ~ "less than or eq to $1",
  price >1 & price <=2 ~ "between $1 and $2",
  price > 2 ~ "more than $2"
)) %>% relocate(group)

produce2 %>% group_by(group) %>% summarise(total_spent = sum(total_cost))
```

```{r}
students = tibble(
  name = c("A", "B", "C", "D", "E", "F"),
  age = 17:22
) %>% 
  slice_sample(n = 6)

students %>% mutate(legal_status = case_when(
  age <= 17 ~ "Can't vote or drink",
  age >= 18 & age < 21 ~ "Can vote but can't drink",
  age >= 21 ~ "Can vote and drink"
))
```






