---
title: "Homework 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE, error = TRUE,
                      fig.height = 3)
library(tidyverse)
```

# Preliminaries

- This file should be in `STAT240/homework/hw05` on your local computer.
- Download `Cameron_Lectures_SP24.csv` to `STAT240/data`.


# Problem 1

In a certain soccer league, there are a total of 1-7 goals per game, with most games having 2 or fewer total goals.

The distribution of the number of goals is shown below, with the probability of four total goals being scored removed.

```{r}
goal_vals <- 1:7
goal_probs <- c(0.335, 0.236, 0.231, NA, 
                0.066, 0.018, 0.012)

goals <- tibble(goal_vals, goal_probs)

goal_distribution <- ggplot(goals, aes(x = goal_vals, y = goal_probs)) +
  geom_segment(aes(xend = goal_vals, yend = 0), color = "blue", size=2) +
  geom_hline(yintercept = 0) +
  scale_x_continuous(breaks = goal_vals) +
  ylab("P(X=x)") +
  ggtitle("Distribution of Total Goals per Game") +
  
  annotate("text", x = 4, y = 0.1, label = "?", size = 15, color = "blue")

goal_distribution
```

**Write code** with either base R or dplyr to manually compute the value of the missing $P(\text{Goals} = 4)$. 

**Save that single value** as `p4`.

On a separate line, let the value of `p4` be **printed as output** in your knitted file.

```{r}
# Write your code here!

```


# Problem 2

The following code inserts the missing `p4` into the missing slot.

```{r}
goal_vals <- 1:7
goal_probs <- c(0.335, 0.236, 0.231, p4, 
                0.066, 0.018, 0.012)
```

**a)** What is the mean of this distribution?

```{r}
# Write your code here!

```

**b)** What is the variance of this distribution?

```{r}
# Write your code here!

```

**c)** What is the standard deviation of this distribution?

```{r}
# Write your code here!

```


# Problem 3

For each of the real life scenarios described below, state whether you would treat the random variable as **discrete** or **continuous** and briefly explain why.

The random process for each of these random variables is: **Your professor holds a STAT 240 lecture.** We have run this "random process" 15+ times already and could have collected 15+ different data points of these random variables!

**a)** The **number** of students who attend a given lecture.

**b)** The **average age** of the students who attend a given lecture. (Treat the age of an *individual student* as a whole number of years, e.g. 18, 19, 20, 21.)

**c)** The **length**, in seconds, of the lecture recording for a specific lecture. *Pretend you only know the length to the nearest round second.*


# Problem 4

The `lectures` dataset contains information on the length *in decimal minutes* (not seconds) of Cameron's **Spring 2024** (last semester!) STAT 240 lectures, and information on the outfits he wore.

```{r}
lectures = read_csv("../../data/Cameron_Lectures_SP24.csv")
```

**Create a histogram** of `total_minutes`.

Incorporate the following customizations:

- The **bins** should be of width 1, and have boundaries on each of the whole numbers. E.g., there should be a bin from 47 to 48, 48 to 49, et cetera.

- Use a consistent color scheme for the interior and exterior of each bar that is the same for every bar.

- If ggplot does not do it by default, make sure the x axis labels are each of the whole numbers 47 through 52.


# Problem 5

Our goal in this problem is to calculate the percentages associated with each of the five bars in the above histogram.

First, create a new variable called `base_minutes` by "truncating" `total_minutes` by using the `floor()` function. *This is effectively just cutting the decimal off, which is different than rounding; 49.4, 49.9, 50.1, 50.8 will become 49, 49, 50, 50.*

Then, create a dataframe with one row for each different value of `base_minutes`, and a column `n` for how many times that value of `base_minutes` occurs. 

Finally, create a `percent` variable by dividing `n` by the total number of lectures.

Let this dataframe **print as output**, no need to save it in a new variable name.

*You should determine that 55.3% of Cameron's Spring 2024 lectures were 50 minutes long (plus extra seconds we do not care about).*


# Problem 6

**a)**
Consider an upcoming, future lecture of Cameron's that has not happened yet - for example, the lecture that will occur on Wednesday, December 4th, 2024.

**Is it valid to say that the future lecture has a 55.3% chance of being 50 minutes + left-over seconds long?** Why or why not?

**b)**
Consider selecting a random lecture from Cameron's set of existing Spring 2024 recordings. 

**Is it valid to say that the randomly selected lecture has a 55.3% chance of being 50 minutes + left-over seconds long?** Why or why not?


# Problem 7

Let's consider modeling the duration of Cameron's upcoming Fall 2024 lectures as a **continuous** variable instead of a discrete one, with the bell-curve model given in `hw05p7a.png`.  According to the Spring 2024 data, this model seems reasonable.

Consider a new data point; Cameron's Fall 2024 lecture on Monday, September 16th, is recorded as being **42 minutes long.**

According to our above model, the probability of a lecture being 42 minutes or shorter is extremely, extremely low.  The probability is visualized in `hw05p7b.png` and the following code computes this probability for you.

```{r}
mu = mean(lectures$total_minutes)
sigma = sd(lectures$total_minutes)

# Calculation of the shaded left-tail probability
pnorm(42, mu, sigma)
```

According to our model, such a short lecture should be extremely unlikely, and yet it happened.

**Does our original model still seem reasonable?** Briefly explain your reasoning.

*Note: This question is about the ability to construct an argument, either yes or no could be justified!*


# Problem 8

**a)**

**Create and print** a table of how often Cameron wore each `top_color` in `lectures`, with percentages included like Problem 5.

*You should find that Cameron wore a navy colored shirt in 36.8% of SP24 lectures.*

**b)**

**Is it valid to guess that approximately 36.8% of all Cameron's shirts he owns are navy-colored?** Briefly explain your reasoning.








