---
title: "week02-in_class_codes"
output: html_document
date: "2025-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, warning = TRUE)
```

# Sec 004

## Monday 1/27

-   Variables:

```{r}
# Numeric variables

a <- 1
A <- 1
a = 1

a1 <- 10
a 1 <- 10 # error because we can't have space in variable names
a_1 <- -10
c <- pi
d <- Inf
e <- -Inf

1a <- 0 # error because can't use number infront of a variable name

# Character variable

my_name <- Sahifa
my_name <- "Sahifa"
sentences <- "it is possible to have sentences as variables in R"

sentences

number <- "2"

number + 3

# Logical variables

x <- TRUE
y <- FALSE
v <- NA

```

-   To know the type of variables: 'class()'

```{r}
class(c)
class(my_name)
class(x)
```

```{r}
x+1
my_name + 1
```

```{r}
X <- 1:100
y <- 1:100
plot(X, y)

class(X)
```
- Operators:

```{r}
# mathematical operators: work with numeric and logical variables except NA

2 + 3 # addition

2 - 3 # subtraction

2 * 3 # multiplication

2 / 3 # division

2 ** 3 # raised to
2 ^ 3 # raised to

TRUE + T
T + F

"Sahifa" + "banana"
cat("Sahifa", "banana")
```

```{r}
# Logical Operators: work with numeric, logical and character except NA

2 < 5
2 <= 2
2 >= 3
3 > 2

2 = 2
```

```{r}
2 == 2
b = T
b == T

2 != 3
2 != 2
T != F

name <- "Sahifa"
name == "Sahifa"
name == "Sahifa "
```

```{r}
# logical operator for 'and' is '&'

T & T
T & F
F & T
F & F
```
```{r}
# the logical operator for 'or' is '|'

T | T 
T | F
F | T 
F | F
```

```{r}
a <- 2
c <- -10

(a < 3) & (c < 0)
```


```{r}
2 + NA
(2 + 2 + 3 + 5) / NA
missing_value <- NA 

missing_value == NA
is.na(missing_value)
!is.na(missing_value)
```

- Function: `function_name(argument1, argument2)`

```{r}
is.na(b)
sqrt(16)
sqrt(16, 2)
?sqrt
sum(1,2,3,4,5)
?sum
sum(1,2,3,4,5,NA, na.rm = T)
mean(1,2,3,4,5)
```

- Vectors: list of objects that get stored under one name

- using the function `c()`, combines all the list of objects into a vector

```{r}
first_5_even_num <- c(2, 4, 6, 8, 10)
class(first_5_even_num)

fruits <- c("apple", "banana", "cherry")

random <- c(2, 1000, TRUE, "Sahifa")

random2 <- c(2, 1, 1000, TRUE, F)

random3 <- c(T, F, 30, NA)

is.na(random3)
```

# Friday 1/31

- using the function `seq()`

```{r}
?seq

seq(0,10)

seq(0, 10, 3)

seq(by = 3, to = 10, from = 0)

list_0_1 <- seq(0, 1, length.out = 20)

list_0_1

seq(0, 1, length.out = 10)

seq(from = 10, to = 1)

seq(1, 10, -1)

seq(10, 1, -2)
```

- using `:`

```{r}
1:10
0.5:20
weird <- pi:21
21: pi
```



```{r}
# calling obejcts from a vector using the index/placeholder

weird

weird[2]

fruits

fruits[3]

weird[1, 5, 9, 13, 17]

weird[c(1, 5, 9, 13, 17)]
```



```{r}
# adding objects/elements to existing vectors

fruits

fruits <- c(fruits, "dragonfruit")

fruits

fruits[5] <- "elderberry"

fruits[c(6,7)] <- c("fig", "grape")

fruits
```

```{r}
# replacing objects inside vector using placeholder

fruits[2] <- "blueberries"

fruits
```

- Operations on a vector

```{r}
# the vector need to be numeric or logical
# they have to be of the same length

firstThreeNum <- 1:3
nextThreeNum <- 4:6

firstThreeNum + nextThreeNum

length(firstThreeNum)
length(nextThreeNum)

firstThreeNum^2

log(firstThreeNum)

sqrt(firstThreeNum)

mean(firstThreeNum)
```


```{r}
length(list_0_1)

firstThreeNum + list_0_1

new <- seq(2, 20, length.out = 12)

new

firstThreeNum

new + firstThreeNum
```

```{r}
# logical operator
new

new == 5

new == 20

20 %in% new

5 %in% new 
```

```{r}
sum(seq(1, 99, by = 2))
```


- Dataframes

- basic R function `data.frame()`
- we will use the function `tibble()` from the package `tidyverse`

```{r}
library(tidyverse)
```


```{r}
?tibble

data <- tibble(num = 1:10, "even or odd" = c("odd", "even", "odd", "even", "odd", "even", "odd", "even", "odd", "even"), isItPrime = num %in% c(2, 3, 5, 7))
```

## Monday 2/3

- Callin/loading datasets from the desktop onto RStudio we will use the function `read_csv("path of the file on the computer")` from the package `tidyverse`

```{r}
mendota <- read_csv("../../data/lake-mendota-winters-2024.csv") # this works only if you have your directory setup in the right way

mendota <- read_csv("~/Desktop/STAT240/data/lake-mendota-winters-2024.csv") # this is for MAC users. for windows use getwd() then copy everything from the start till OneDrive paste it in place of `~/`
```

- Functions that can be used to visualize or understand the data: View(name of the data), head(name of the data), etc

- extracting columns from data as vectors using `$` operator

```{r}
mendota$year1

mendota$duration
```

-  the placeholder can be used to extract columns or rows or cells from a dataset. It extracts the information as a dataset

```{r}
year1 = mendota$year1

year_1 <- mendota[ , 2]
```










