---
title: "week02-in_class_codes"
output: html_document
date: "2025-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, warning = TRUE)
```



# Sec 003

## Monday 1/27

Basic Object:

-  Variables:

```{r}
# numeric variables

a <- 1 # you can also assign values using =
A <- 1

a <- 2 # redefining variables
a <- A+2 # you can use other variables to define your variable

b <- -1
b1 <- 2
b 1 <- 2 # no space in variable name
b_1 <- 2
b.1 <- 3

# Character variables

food <- cheese # this is a character but since it's not inside of quotation marks, it gets recognized as a variable name instead
food <- "cheese" #you need to use quotation marks around the value
name <- "My name is Sahifa"
number <- "2"
number + 2

# Logical variables

x <- TRUE
y <- FALSE 
NA
```

-   How to know the type of variables?

Use function: `class()`

```{r}
x+1
class(b_1)
class(number)
class(x)
class(TRUE)
```

```{r}
x <- 1:4/2
class(x)
x[2]
```


# Wednesday 1/29

- Operators

```{r}
# mathematical operators: only work with numerical object or logical objects

2 + 3 # addition

2 - 3 # subtraction

2 * 3 # multiplication

2 / 3 # division

2 ^ 3 # raised to
2 ** 3 # raised to

T + T
T - F

T/F

name + food
```

```{r}
# Logical Operator

2 < 5
5 <= 5 
5 > 6
5 >= 5

5 = 5
5 = -1

5 == 5
5 == 2

5 != 1

food == "cheese"
food == "cheese "
food == "cheese_"

T == T
```


- Disclaimer: No operators work with NAs

```{r}
2+2+3+T
2+2+3+NA
b <- NA
b == NA
is.na(b)
```



```{r}
# logical operator for 'and' is '&"

T & T
T & F
F & T
F & F

```

```{r}
# logical operator for the 'or' is '|'

T | T
T | F
F | T 
F | F
```

```{r}
a <- 5
b <- -5
(a > 3) & (b < 0)
```

- Functions:

```{r}
sqrt(4)
sqrt(4, 16)
class(b)

sum(1, 2, 3, 4, 5)
?sum
sum(1, 2, 3, 4, 5, NA)
sum(1, 2, 3, 4, 5, NA, na.rm = TRUE)
```


- Vector: is a list of objects that gets saved under a variable name

Ways of creating a vector:

- using the function `c()`, combine

```{r}
even <- c(2, 4, 6, 8, 10)
class(even)

fruit <- c("apple", "banana", "cantalope")
fruit <- c('apple', 'banana', 'cantalope')

random <- c(2, 3, 1000, TRUE, "name", "2")
random

random2 <- c(2, 3, TRUE, FALSE, 1000)
random2
```

## Friday 1/31

- sequences: using function `seq()` or using colon between two numbers `:`

```{r}
?seq
seq(1, 100)
1:100

seq(1, 100, 2)

seq(from = 1, to = 10, by = 1)

list_0_1 <- seq(0, 1, length.out = 100)

reverse <- seq(10, 1, by = -1)

reverse

seq(0, 1, 0.1)
```

- using `:`

```{r}
1:10

10:1

pi:21

seq(pi, 21, by = pi)
```



```{r}
# calling objects using their placeholders/indexes

fruit[3]

list_0_1[1,20, 40, 50]

list_0_1[c(1,20, 40, 50)]
```

```{r}
# new objects or replacing objects inside of a vectors

fruit

fruit <- c(fruit, "mango")

fruit

fruit[5] = "watermelon"

fruit[2] = "strawberry"

fruit
```

- Mathematical operations:

```{r}
# the vectors ahve to be numeric or logical
# they have to be of the same length

firstThreeNum <- 1:3
nextThreeNum <- 4:6

firstThreeNum+nextThreeNum

firstThreeNum ^ nextThreeNum

firstThreeNum^2

log(even)

sqrt(even)

mean(even)

length(even)
```

```{r}
# mathematical operations with uneven lengths of vectors

even

list <- 1:20

list

even + list

even

firstThreeNum

even + firstThreeNum
```


```{r}
# checking for an object inside of a vector

fruit

fruit == "kiwi"

"kiwi" %in% fruit

0.2 %in% list_0_1 
```


```{r}
sum(seq(1, 99, 2))
```


- Dataframe:

We create data frames using the function `tibble()` from the package `tidyverse`.

```{r}
library(tidyverse)
```


```{r}
?tibble

data <- tibble(num = 1:10, "even or odd" = c("odd", "even", "odd", "even", "odd", "even", "odd", "even", "odd", "even"), "isItPrime" = num %in% c(2, 3, 5, 7))
```

##  Monday 2/3

- Calling/loading datasets from the computer onto RStudio using the function `read_csv(path to dataset file on the computer)` from the `tidyverse` package

```{r}
mendota <- read_csv("../../data/lake-mendota-winters-2024.csv") # set the correct working directory

mendota <- read_csv("~/Desktop/STAT240/data/lake-mendota-winters-2024.csv") # thsi is going to work for any working directory. for windows user, instead of ~/ use the initial part of getwd output till OneDrive
```

- Functions that can be used to visualize or understand the data: View(name of the data), head(name of the data), etc

- extracting columns from data as vectors using `$` operator
```{r}
mendota$winter
```


```{r}
# the placeholder can be used to extract columns or rows or cells from a dataset. It extracts the information as a dataset

mendota[ , 1]
mendota[169, ]
mendota[169,2]
```







